---
title: "Map demo for RUG"
subtitle: "By Gina Contolini. Gina.Contolini@noaa.gov"
author: "Greater Farallones Association and Greater Farallones and Cordell Bank National Marine Sanctuaries"
date: "Updated on `r Sys.Date()`"
output: 
  html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(tmap)
library(leaflet)
library(ggmap)
library(ggspatial)
```

# Tmap package
Thematic maps are geographical maps in which spatial data distributions are visualized. This package offers a flexible, layer-based, and easy-to-use approach to create thematic maps, such as choropleths and bubble maps.
The following example is based on a demo from [thinking-spatial.org](https://thinking-spatial.org/courses/angewandte_geodatenverarbeitung/kurs04/).  
```{r tmap quick demo}
# the following code is based on this online demo: thinking-spatial.org/courses/angewandte_geodatenverarbeitung/kurs04 
# load dataset
data(World) 
# look at data
# glimpse(World)

# make a static map with that data
tm_shape(World) +
  tm_polygons("HPI")

# make an interactive map with that data
tmap_mode('view') # change the mode to interactive view
tm_shape(World) +
  tm_polygons('HPI')
# cool beans
```

# Leaflet package
Create and customize interactive maps using the "Leaflet" JavaScript library and the "htmlwidgets" package. These maps may be used directly from the R console, from RStudio, in Shiny applications, and R Markdown documents.
The following examples are based on a demo from [jla-data.net](https://www.jla-data.net/eng/leaflet-in-r-tips-and-tricks/).

```{r leaflet quick demo}
# the following code is based on this online demo: https://www.jla-data.net/eng/leaflet-in-r-tips-and-tricks/ 

# set up a data frame of points
df <- data.frame(name = c("Kramářova vila", "Pražský hrad", "Strakova akademie", "Nejvyšší soud", "Ústavní soud", "Sněmovna", "Senát"),
          branch = c("executive", "executive", "executive", "judiciary", "judiciary", "legislature", "legislature"),
          lat = c(14.4104392, 14.3990089, 14.4117831, 16.6021958, 16.6044039, 14.4039458, 14.4053489),
          lon = c(50.0933681, 50.0895897, 50.0920997, 49.2051925, 49.1977642, 50.0891494, 50.0900269)) %>% 
  sf::st_as_sf(coords = c("lat", "lon"), crs = 4326) # transform the data frame from plain vanilla one to spatial. # WGS84 is a sensible default

# prepare leaflet plot
leaflet(data = df) %>% # creates leaflet object
  addTiles() %>% # adds basemap
  addMarkers() # adds data layers, in this case markers

# This makes a cool map you can zoom into. Here's a new example with kelp team data.

# load urchin removal points
  urchin_removals <- read.csv('Purple urchin landings datasheet - removal_coords.csv') # read.csv might be different than read_csv?

# show all markers
leaflet(data = urchin_removals) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addMarkers() 

# Code markers by vessel, and add date on hover (I actually don't know why it's adding the date, but it's cool!)
# create aesthetic with vessel colors
vessel_colors <- leaflet::colorFactor(palette = c('Guardian' = 'red',
                                  'Omega' = 'goldenrod', 
                                  'Janae' = 'steelblue'), 
                      domain = urchin_removals$vessel)
# create the map
leaflet(data = urchin_removals) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircleMarkers(radius = 5, popup = ~date, color = vessel_colors(urchin_removals$vessel), fillOpacity = 0.7, stroke = FALSE) %>%
  leaflet::addLegend(position = "topright",
            values = ~vessel, # data frame column for legend
            opacity = .7, # alpha of the legend
            pal = vessel_colors, # palette made earlier
            title = "Vessel") # legend title

# Cluster markers
leaflet(data = urchin_removals) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircleMarkers(radius = 10, popup = ~date, color = vessel_colors(urchin_removals$vessel), stroke = FALSE, clusterOptions = markerClusterOptions())
```

# SF package
Support for simple features, a standardized way to encode spatial vector data. Binds to "GDAL" for reading and writing data, to "GEOS" for geometrical operations, and to "PROJ" for projection conversions and datum transformations. Uses by default the "s2" package for spherical geometry operations on ellipsoidal (lon/lat) coordinates.
I have no example using SF because it wasn't compatible with my version of R, but I can help you find an online demo if you're interested.

# ggmap package
A collection of functions to visualize spatial data and models on top of static maps from various online sources (e.g., Google Maps and Stamen Maps). It includes tools common to those tasks, including functions for geolocation and routing. The following example is based on a demo from [thinking-spatial.org](https://thinking-spatial.org/courses/angewandte_geodatenverarbeitung/kurs05/).  

```{r ggmap demo with kelp data}
# The following code is based on a demo from https://thinking-spatial.org/courses/angewandte_geodatenverarbeitung/kurs05/
# load coordinates you want to plot
removal_coords <- read_csv('Purple urchin landings datasheet - removal_coords.csv')
# make date a date class so it plots continuous not categorical
removal_coords$date = dmy(removal_coords$date)

# must get API key first for some reason. Do the following once per R session:
ggmap::register_stadiamaps('8d134a7f-55f5-4eac-a02c-eecd8c3d4784') #here's where I got that API key: https://client.stadiamaps.com/dashboard/#/property/19840/. Eventually it will probably expire and I'll have to get a new API key.

## North arrow and scale bar features demonstrated here: https://stackoverflow.com/questions/77437044/add-scalebar-and-north-arrow-in-a-ggmap-without-using-ggsn

# load coordinates to plot
#### Create the map! ####
frc_box = c(-123.252, 38.508, -123.24, 38.515) # This is the bounding box. c(lowerleftlon, lowerleftlat, upperrigthlon, upperrightlat)

ggmap(get_stadiamap(bbox = frc_box, zoom = 16)) + #higher zoom increases resolution
  geom_point(aes(x = lon, y = lat), alpha = .5, data = removal_coords,  size = 2.2) +
  ggtitle('Purple sea urchin removal locations at Fort Ross Cove in 2023') +
  labs(x = 'Longitude', y = 'Latitude') +
  ggspatial::annotation_north_arrow(location = "tr", style = ggspatial::north_arrow_minimal()) + #north arrow
  annotation_scale(location = "bl", width_hint = 0.1) + # scale bar
  coord_sf(crs = 4326) # assuming this is the crs.

```
